---
title: "Analysis"
author: "Nagaprasad Rudrapatna" 
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      warning = F, 
                      message = F)
```

```{r}
library(tidyverse)
library(knitr)
library(lme4)
library(broom.mixed)
library(viridis)
library(skimr)
```

```{r}
lemurs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')
```

```{r}
ggplot(data = lemurs, aes(x = current_resident, y = weight_g)) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Weight (in grams)", 
       subtitle = "by residency status")

lemurs %>%
  count(current_resident) %>%
  mutate(p = n / sum(n))
```

no clear relationship between `current_resident` and `weight_g`

```{r}
ggplot(data = lemurs, aes(x = avg_daily_wt_change_g))  + 
  geom_histogram(fill = "steelblue", color = "black") + 
  labs(title = "Avg Daily Change in Weight (in grams) between two consecutive weights")
```

```{r}
# removing irrelevant variables
lemurs2 <- subset(lemurs, select = -c(dlc_id, stud_book, birth_month, estimated_dob, estimated_concep, concep_month, dam_id, dam_taxon, dam_dob, dam_age_at_concep_y, sire_id, sire_taxon, sire_dob, sire_age_at_concep_y, age_at_death_y, age_of_living_y, age_last_verified_y, age_max_live_or_dead_y, dob_estimated, change_since_prev_wt_g, avg_daily_wt_change_g, r_min_dam_age_at_concep_y, expected_gestation_d, concep_date_if_preg, infant_dob_if_preg, pct_preg_remain_if_preg, current_resident, litter_size, sire_name, month_of_weight, age_at_wt_wk, age_at_wt_mo, age_at_wt_mo_no_dec, days_since_prev_wt, days_before_death, days_before_inf_birth_if_preg, hybrid, expected_gestation))
```

all 323 animals in the final dataset are non-hybrids

```{r}
ggplot(data = lemurs2, aes(x = sex, y = weight_g)) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Weight (in grams)", 
       subtitle = "by sex")

lemurs2 %>%
  count(sex == "ND") %>%
  mutate(p = n / sum(n))
```

- not determined sex: less than 0.018% of animals
- remove them 

```{r}
list <- lemurs2 %>%
  group_by(name) %>% 
  count() %>%
  filter(n <= 2)
nrow(list)
```

- 598 animals whose weight was measured only once or twice
- remove these

```{r}
lemurs3 <- lemurs2 %>%
  filter(sex != "ND")

for (i in 1:nrow(list)) {
  lemurs3 <- lemurs3 %>%
  filter(lemurs3$name != list$name[i])
}
```

```{r}
ggplot(data = lemurs3, aes(x = weight_g))  + 
  geom_histogram(fill = "steelblue", color = "black") + 
  labs(title = "Weight (in grams)")
```

```{r}
ggplot(data = lemurs3, aes(x = taxon, y = weight_g)) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Weight (in grams)", 
       subtitle = "by taxonomic code")
```

highest median weight - Coquerel’s sifaka (PCOQ)
lowest median weight - Gray mouse lemur (MMUR)

small lemurs cause issues with the distribution of response; also taxon has too many levels to be included as is in the model as fixed effect

```{r}
lemurs3 %>%
  filter(weight_g >= 7500)

# exclude extremely high weight measurements

lemurs3 <- lemurs3 %>%
  filter(weight_g < 7500)

lemurs3 %>%
  group_by(taxon) %>%
  summarise(median = median(weight_g), mean = mean(weight_g), n = n()) %>%
  arrange(desc(mean))
```

limit it to the 3 heaviest lemur taxons on average
- PCOQ, VRUB, and VVV

```{r}
lemurs3 <- lemurs3 %>%
  filter(taxon == "PCOQ" | taxon == "VRUB" | taxon == "VVV")
```

```{r}
lemurs3 %>%
  count(sex) %>%
  mutate(p = n / sum(n))
```

- dataset has roughly equal proportions of males and females

** Analysis applies to heavy lemurs

```{r}
ggplot(data = lemurs3, aes(x = weight_g))  + 
  geom_histogram(fill = "steelblue", color = "black") + 
  labs(title = "Weight (in grams)")

ggplot(data = lemurs3, aes(x = log(weight_g)))  + 
  geom_histogram(fill = "steelblue", color = "black") + 
  labs(title = "Log Weight (in grams)")

ggplot(data = lemurs3, aes(x = log(weight_g)))  + 
  geom_histogram(fill = "steelblue", color = "black") + 
  labs(title = "Log Weight (in grams)") +
  facet_wrap(~ taxon)
```

spike at low weights (IJ)
right-skewed 
unimodal

log-transform response:
left-skewed, unimodal

```{r}
lemurs3 %>%
  group_by(taxon) %>%
  count(age_category)
```

most IJ are of PCOQ taxon

```{r}
ggplot(data = lemurs3, aes(x = taxon, y = log(weight_g))) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Log Weight (in grams)", 
       subtitle = "by taxonomic code")
```

```{r}
ggplot(data = lemurs3, aes(x = preg_status, y = log(weight_g))) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Log Weight (in grams)", 
       subtitle = "by pregnancy status")

lemurs3 %>%
  filter(sex == "F") %>%
  count(preg_status) %>%
  mutate(p = n / sum(n))

lemurs3 %>%
  filter(sex == "F") %>%
  group_by(preg_status) %>%
  summarise(median = median(weight_g))
```

- roughly 7.9% of females were pregnant during a weight check
- pregnant animals have higher median weight

```{r}
ggplot(data = lemurs3, aes(x = age_category, y = log(weight_g))) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Log Weight (in grams)", 
       subtitle = "by age category")

lemurs3 %>%
  count(age_category) %>%
  mutate(p = n / sum(n))

lemurs3 %>%
  group_by(sex) %>%
  count(age_category) %>%
  mutate(p = n / sum(n))
```

- roughly equal proportions of males and females in each age category

~ 52% adult
~ 35% IJ
~ 13% young adult

- young adults have similar median weight to adults, IJ have lowest median weight

```{r}
lemurs3$birth_type <- as.factor(lemurs3$birth_type)

lemurs3 %>%
  group_by(birth_type) %>%
  summarise(median = median(weight_g), n = n())

# remove unknown birth type

lemurs3 <- lemurs3 %>%
  filter(birth_type != "Unk")

ggplot(data = lemurs3, aes(x = birth_type, y = log(weight_g))) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Log Weight (in grams)", 
       subtitle = "by birth type")
```

wild birth has slightly higher median log-weight

```{r}
lemurs3$birth_institution <- ifelse(lemurs3$birth_institution == "Duke Lemur Center", "Duke Lemur Center", "Other")

lemurs3$birth_institution <- as.factor(lemurs3$birth_institution)

lemurs3 %>%
  ggplot(aes(x = birth_institution, y = log(weight_g))) + 
  geom_boxplot(fill = "steelblue", color = "black")  +
  labs(title = "Log Weight, in grams", 
       subtitle = "by birth institution")

lemurs3 %>%
  group_by(birth_institution) %>%
  count() 
lemurs3 %>%
  group_by(birth_institution) %>%
  summarise(median = median(weight_g))
```

since the overwhelming majority (about 85.5%) of lemurs were born at DLC, we can simply consider DLC or not

lemurs not born at the DLC has higher median weight

```{r}
# if n_known_offspring is missing, replace with 0

lemurs3$n_known_offspring <- ifelse(is.na(lemurs3$n_known_offspring) == T, 0, lemurs3$n_known_offspring)

ggplot(data = lemurs3, aes(x = n_known_offspring, y = log(weight_g))) +
  geom_point() +
  labs(x = "Number of known offspring",
       y = "Log Weight in grams", 
       title = "Log Weight as a function of known offspring")

lemurs3 %>%
  group_by(infant_lit_sz_if_preg) %>%
  count()
```

no clear relationship with `n_known_offspring`

```{r}
lemurs3$infant_lit_sz_if_preg <- ifelse(is.na(lemurs3$infant_lit_sz_if_preg) == T, 0, lemurs3$infant_lit_sz_if_preg)

ggplot(data = lemurs3, aes(x = infant_lit_sz_if_preg, y = log(weight_g))) +
  geom_point() +
  labs(x = "Size of Litter",
       y = "Log Weight in grams", 
       title = "Log Weight as a function of litter size when giving birth")
```

no relationship with litter size

```{r}
lemurs3 %>% skim() %>% select(skim_variable, n_missing, complete_rate)
```

```{r}
lemurs3$log_weight <- log(lemurs3$weight_g)
```

```{r}
# thus, we cannot have any age = 0 (weight checks on DOB)
# use days because for very young lemurs, years/months not interpretable

lemurs3 <- lemurs3 %>%
  filter(age_at_wt_d > 0)

lemurs3 %>%
  distinct(name) %>%
  count()
```

323 unique animals

```{r}
set.seed(7)

# log-transform age to get linearity

lemurs3$log_age_at_wt_d <- log(lemurs3$age_at_wt_d)

# get sample of 24 animals
sample_lem <- lemurs3 %>%
  distinct(name) %>%
  sample_n(24) %>% pull()

# get data for those animals
sample_data <- lemurs3 %>%
  filter(name %in% sample_lem)
```

```{r}
## Make a lattice plot of the weights over time for 24 randomly selected animals.

ggplot(data = sample_data, aes(x = age_at_wt_d, y = log_weight)) +
  geom_point() +
  geom_line() +
  labs(x = "Log of age when weight was taken, in days",
       y = "Log Weight in grams", 
       title = "Weight over time for lemurs") +
  facet_wrap(~ name)
```

```{r}
## Make a lattice plot of the weights over time for 24 randomly selected animals.

ggplot(data = sample_data, aes(x = log_age_at_wt_d, y = log_weight)) +
  geom_point() +
  geom_line() +
  labs(x = "Log of age when weight was taken, in days",
       y = "Log Weight in grams", 
       title = "Weight over time for lemurs") +
  facet_wrap(~ name)
```

```{r}
## spaghetti plot

ggplot(data = lemurs3, aes(x = age_at_wt_d, y = weight_g)) + 
  geom_line(aes(group = name), color = "light gray") + 
  geom_smooth(color = "black", size = 1) + 
  labs(x = "Log of age when weight was taken, in days",
       y = "Log Weight in grams", 
       title = "Weight over time for lemurs")
```

```{r}
## spaghetti plot

ggplot(data = lemurs3, aes(x = log_age_at_wt_d, y = log_weight)) + 
  geom_line(aes(group = name), color = "light gray") + 
  geom_smooth(color = "black", size = 1) + 
  labs(x = "Log of age when weight was taken, in days",
       y = "Log Weight in grams", 
       title = "Weight over time for lemurs")
```

```{r}
# spaghetti plot by sex
ggplot(data = lemurs3, aes(x = log_age_at_wt_d, y = log_weight)) + 
  geom_line(aes(group = name), color = "light gray") + 
  geom_smooth(color = "black", size = 1) + 
  labs(x = "Log of age when weight was taken, in days",
       y = "Log Weight in grams", 
       title = "Weight over time for lemurs") +
  facet_wrap(~ sex)
```

## Unconditional means model 

```{r}
# Fit the unconditional means model
model1 <- lmer(log_weight ~ 1 + (1 | name), data = lemurs3, REML = F)
tidy(model1)
```

```{r}
# Calculate the intraclass correlation
0.7147289^2 / (0.7147289^2 + 0.7801945^2)
```

p-hat = 0.456
~ 45.6% of the variability in log-weight measurements can be attributed to differences between lemurs (lemur-to-lemur variability). The average correlation between any two log-weight measurements for the same animal is about 0.456.

About 54.4% of the variability in log-weight measurements can be attributed to changes over time

```{r}
# Fit the unconditional means model with parent - only mother since sire data is missing
model1a <- lmer(log_weight ~ 1 + (1 | name) + (1 | dam_name), data = lemurs3, REML = F)
tidy(model1a)
```

```{r}
sigma_um1 <- 0.7801945^2
sigma_um2 <- 0.8337123^2
p_r2a <- (sigma_um1 - sigma_um2) / sigma_um1
p_r2a
```

~ 14.2% increase in the residual variance due to inclusion of dam as a random effect

## Unconditional growth model

```{r}
# Fit the unconditional growth model
model2 <- lmer(log_weight ~ log_age_at_wt_d + (log_age_at_wt_d | name), data = lemurs3, REML = F)
tidy(model2)
```

Calculate the $Pseudo R^2$ to estimate the change of within-animal variance between the unconditional means and unconditional growth models.

```{r}
# Pseudo R^2
sigma_um <- 0.7801945^2
sigma_ug <- 0.2421549^2
p_r2 <- (sigma_um - sigma_ug) / sigma_um
p_r2
```

~ 90.4% decrease in the residual variance due to inclusion of `log_age_at_wt_d` (age at time of weight check)

# Adding Level One variables

```{r}
lemurs3$age_category <- as.factor(lemurs3$age_category)
lemurs3$preg_status <- as.factor(lemurs3$preg_status)
```

```{r}
model3 <- lmer(log_weight ~ log_age_at_wt_d + age_category + (log_age_at_wt_d | name), data = lemurs3, REML = F)
tidy(model3)
```

```{r}
# comparing models 2 and 3 using LRT with chi-squared
anova(model3, model2, test = "Chisq")
```

p-value: < 2.2e-16
model 3 > model 2 (include age_category)

also, model 3 has lower AIC/BIC

```{r}
model4 <- lmer(log_weight ~ log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d | name), data = lemurs3, REML = F)
tidy(model4)
```

```{r}
# comparing models 3 and 4 using LRT with chi-squared
anova(model4, model3, test = "Chisq")
```

p-value: 2.381e-14
model 4 > model 3 (include `age_category`, `preg_status`)

model 4 has lower AIC/BIC

Level One covariates: `log_age_at_wt_d`, `age_category`, and `preg_status`

## Adding Random Effects

```{r}
model5 <- lmer(log_weight ~ log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model5)
```

```{r}
# comparing models 4 and 5
anova(model5, model4, test = "Chisq")
```

parametric bootstrap takes too long but p-value from LRT is already very small
p-value: < 2.2e-16
model 5 > model 4 
model 5 has lower AIC/BIC

```{r}
model6 <- lmer(log_weight ~ log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category + preg_status | name), data = lemurs3, REML = F)
tidy(model6)
```

```{r}
# comparing models 5 and 6
anova(model6, model5, test = "Chisq")
```

model 6 has lower AIC/BIC but the difference is very small (AIC diff: 50.2; BIC diff: 11.9)

does not justify the 5 additional variance components, so despite the LRT result, stick with model 5

```{r}
model7 <- lmer(log_weight ~ log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d | name) + (age_category | name), data = lemurs3, REML = F)
tidy(model7)
```

```{r}
anova(model5, model7, test = "Chisq")
```
stick with model 5 since it has lower AIC/BIC

```{r}
tidy(model5)
```

## Adding Level Two (lemur-specific) covariates

```{r}
lemurs3$taxon <- as.factor(lemurs3$taxon)
lemurs3$sex <- as.factor(lemurs3$sex)
```

Level One: 

$$
\log(Y_{ij}) = a_i + b_i \log(age)_{ij} + c_i ageCategoryIJ_{ij} + d_i  ageCategoryYoungAdult_{ij} + \epsilon_{ij}
$$
---------
u_i - intercept
v_i - log(age)
y_i - age_categoryIJ
z_i - age_categoryYoungAdult

Level 2 potential covariates: sex, taxon, birth_institution, birth_type, n_known_offspring, infant_lit_sz_if_preg

only want categorical-quantitative interactions

Level Two:

intercept: taxon, sex, birth_type, infant_lit_sz_if_preg

log-age: taxon, birth_type

age_categoryIJ: infant_lit_sz_if_preg

age_categoryYoungAdult: birth_type, infant_lit_sz_if_preg

$$
\begin{aligned}
a_i &= \alpha_0 + \alpha_1 taxonVRUB_{i} + \alpha_2 taxonVVV_{i} + \alpha_3 sexM_{i} + \alpha_4 birthTypeWB_{i} + \alpha_5 infantLitterSize_{i} + u_i \\
b_i &= \beta_0 + \beta_1 taxonVRUB_{i} + \beta_2 taxonVVV_{i} + \beta_3 birthTypeWB_{i} + v_i \\
c_i &= \gamma_0 + \gamma_1 infantLitterSize_{i} + y_i \\
d_i &= \psi_0 + \psi_1 birthTypeWB_{i} + \psi_2 infantLitterSize_{i} + z_i
\end{aligned}
$$

Composite:

$$
\begin{aligned}
\log(Y_{ij}) &= \alpha_0 + \alpha_1 taxonVRUB_{i} + \alpha_2 taxonVVV_{i} + \alpha_3 sexM_{i} + \alpha_4 birthTypeWB_{i} + \alpha_5 infantLitterSize_{i} \\ 
&+ [\beta_0 + \beta_1 taxonVRUB_{i} + \beta_2 taxonVVV_{i} + \beta_3 birthTypeWB_{i}] \times \log(age)_{ij} \\
&+ [\gamma_0 + \gamma_1 infantLitterSize_{i}] \times ageCategoryIJ_{ij} \\
&+ [\psi_0 + \psi_1 birthTypeWB_{i} + \psi_2 infantLitterSize_{i}] \times ageCategoryYoungAdult_{ij} \\
&+ u_i + v_i \log(age)_{ij} + y_i ageCategoryIJ_{ij} + z_i ageCategoryYoungAdult_{ij} + \epsilon_{ij}
\end{aligned}
$$
preg_status was in final level 1 only model as a fixed effect
preg_status dropped once we added level 2 covariates

infant_lit_sz_if_preg:age_categoryIJ appears insignificant (t-statistic < 2) but the other level of the interaction is significant, so we keep it

# Intercept

```{r}
# Fit models with animal-specific covariates
model8 <- lmer(log_weight ~ taxon + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model8)
```

```{r}
anova(model8, model5, test = "Chisq")
```

p-value: 0.0002466
model8 > model5 (include taxon)
model8 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model9 <- lmer(log_weight ~ taxon + sex + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model9)
```

```{r}
anova(model9, model8, test = "Chisq")
```

p-value: 0.007209
model9 > model8 (include taxon, sex)
model9 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model10 <- lmer(log_weight ~ taxon + sex + birth_institution + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model10)
```

```{r}
anova(model10, model9, test = "Chisq")
```

p-value: 0.5324
model9 > model10 (include taxon, sex)
model9 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model11 <- lmer(log_weight ~ taxon + sex + birth_type + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model11)
```

```{r}
anova(model11, model9, test = "Chisq")
```

p-value: 0.0142
model11 > model9 (include taxon, sex, birth_type)
model11 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model12 <- lmer(log_weight ~ taxon + sex + birth_type + n_known_offspring + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model12)
```

```{r}
anova(model12, model11, test = "Chisq")
```

p-value: 0.1792
model11 > model12 (include taxon, sex, birth_type)
model11 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model13 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model13)
```

```{r}
anova(model13, model11, test = "Chisq")
```

p-value: 0.007209
model13 > model11 (include taxon, sex, birth_type, infant_lit_sz_if_preg)
model13 has lower AIC/BIC

# Log-Age

```{r}
# Fit models with animal-specific covariates
model14 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model14)
```

```{r}
anova(model14, model13, test = "Chisq")
```

p-value: < 2.2e-16
model14 > model13 (include taxon)
model14 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model15 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:sex + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model15)
```

```{r}
anova(model15, model14, test = "Chisq")
```

p-value: 0.4186
model14 > model15 (include taxon)
model14 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model16 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model16)
```

```{r}
anova(model16, model14, test = "Chisq")
```

p-value: < 2.2e-16
model16 > model14 (include taxon, birth_type)
model16 has lower AIC/BIC

# Age_category

```{r}
# Fit models with animal-specific covariates
model17 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:taxon + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model17)
```

```{r}
anova(model17, model16, test = "Chisq")
```

p-value: 0.5035
model16 > model17 
model16 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model18 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:sex + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model18)
```

```{r}
anova(model18, model16, test = "Chisq")
```

p-value: 0.3941
model16 > model18
model16 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model19 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:birth_type + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model19)
```

```{r}
anova(model19, model16, test = "Chisq")
```

p-value: 3.585e-05
model19 > model16 (include birth_type)
model19 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model20 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:birth_type + age_category:infant_lit_sz_if_preg + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model20)
```

```{r}
anova(model20, model19, test = "Chisq")
```

p-value: 0.0422
model20 > model19 (include birth_type, infant_lit_sz_if_preg)
model20 has lower AIC/BIC

# Preg_status

```{r}
# Fit models with animal-specific covariates
model21 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:birth_type + age_category:infant_lit_sz_if_preg + preg_status:taxon + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model21)
```

```{r}
anova(model21, model20, test = "Chisq")
```

p-value: 0.3561
model20 > model21 
model20 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model22 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + preg_status + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:birth_type + age_category:infant_lit_sz_if_preg + preg_status:birth_type + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model22)
```

```{r}
anova(model22, model20, test = "Chisq")
```

p-value: 0.4755
model20 > model22 
model20 has lower AIC/BIC

```{r}
# Fit models with animal-specific covariates
model23 <- lmer(log_weight ~ taxon + sex + birth_type + infant_lit_sz_if_preg + log_age_at_wt_d + age_category + log_age_at_wt_d:taxon + log_age_at_wt_d:birth_type + age_category:birth_type + age_category:infant_lit_sz_if_preg + (log_age_at_wt_d + age_category | name), data = lemurs3, REML = F)
tidy(model23)
```

```{r}
anova(model20, model23, test = "Chisq")
```

p-value: 0.464
model23 > model20
model23 has lower AIC/BIC

- drop preg_status

# Checking assumptions

```{r}
model23 %>%
  augment() %>%
  mutate(.std.resid = residuals(model23, scaled = T))
```

```{r}
# residuals v fitted
plot(model23, main = "Residuals vs. Fitted", xlab = "Fitted Values", ylab = "Residuals")
# looks reasonable - there may be some nonlinearity based on the curvature for low fitted values 

# checking the normality of residuals:
qqnorm(resid(model23), main = "Q-Q Plot for the Residuals")

# checking the normality of the random effects:
qqnorm(ranef(model23)$name$`(Intercept)`, 
       main="Q-Q Plot for the Random Intercept (Name)")
```

